### 2024-02-04 ms/simulations.qmd
added random effect of id to simulations and model

---
title: "Simulating and fitting synthetic data"
format: pdf
editor: visual
---

## Purpose

We are simulating and fitting synthetic data sets to develop and validate the statistical model for later analysis. This will eventually go in the supporting information.

## Hyperparameters

Here we explaining hyperparameter choices.

`mu_intercept`: this is the value of $A$ where $\text{log}(g_text{sw}) = 0$ or $g_text{sw} = 1$, so this should be close to $A_\text{max}$, which typically varies between 10 and 50 $\mu~\text{mol}~\text{m}^{-2}~\text{s}^{-1}$. So let's say mean of 30 (for high light).

`sigma_intercept_rep`: variation in intercept among replicates within accession. Let's say 2?

, 20 / 3)