### 2024-02-26 ms/notes.qmd
profiling and improving performance of stan model

---
title: "Notes"
format: html
editor: visual
execute:
  enabled: true
---

Making some notes on methods that might be useful for writing later

## 2024-02-25

Actually, I did this profiling wrong, but I'm going to try model without autocorrelation anyway to compare.

I profiled the `solanum-aa` code and found that most (90.6%) of the time is spent on the `multi_normal` calculations.

Next, I profiled the same model, ignoring residual autocorrelation. The speed-up is substantial (80\%). The time in this model is pretty evenly divided among various calculations and likelihood.

Note that total runtime was 5871.4 seconds for 400 iterations

| name         | total_time |
|--------------|------------|
| CO2_r likelihood | 297.691 |
| CO2_s likelihood | 238.332 |
| H2O_r likelihood | 304.786 |
| H2O_s likelihood | 306.439 |
| LICOR calcs | 1047.940 |
| priors | 749.983 |  
| regression | 645.282 |

Given the efficiency of model without residual autocorrelatiln, I will try fitting actual data with uncorrelated residuals first. If the residuals are highly correlated, then I will try a combination of thinning or accounting for autocorrelation, perhaps finding a more efficient approach.