### 2024-04-06 ms/test.qmd
working on table of estimated parameters

---
title: Unknown
format:
  pdf:
    keep-tex: true  
    toc: false
header-includes: |
  \usepackage{hyperref}
  \addtokomafont{disposition}{\rmfamily}
  \usepackage{siunitx}
  \newcommand{\aax}{$\mathrm{AA}$}
  \newcommand{\Aamphi}{$A_{\mathrm{amphi}}$}
  \newcommand{\Ahypo}{$A_{\mathrm{hypo}}$}
  \newcommand{\agcurve}{$A \textendash g_\text{sw}$}
  \newcommand{\ca}{$C_\mathrm{a}$}
  \newcommand{\caequals}[1]{$C_\mathrm{a} = \qty{#1}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\cabetween}[2]{#1 < $C_\mathrm{a} < \qty{#2}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\gsw}{$g_\text{sw}$}
  \newcommand{\ppfd}{$\mathrm{PPFD}$}
  \newcommand{\ppfdequals}[1]{$\mathrm{PPFD} = \qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
  \newcommand{\rh}{$\mathrm{RH}$}
  \newcommand{\rhequals}[1]{$\mathrm{RH} = #1\%$}
  \newcommand{\tleafequals}[1]{$T_\mathrm{leaf} = \qty{#1}{\degreeCelsius}$}
mainfont: Times New Roman
bibliography: solanum-aa.bib
csl: https://www.zotero.org/styles/science
---

```{r load-objects, echo=FALSE, eval=TRUE}
library(dplyr)
library(readr)
library(tibble)
```

Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into six sequential steps \autoref{tbl-parameters}.

```{r, parameters, echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-parameters
#| tbl-cap: "caption"

df_parameters = read_csv("../data/parameters1.csv", col_types = "cccc", n_max = 6)

df_parameters |>
  dplyr::select(Parameter = latex, Description = description) |>
  knitr::kable(
    format = "latex",
    align = c("c", "l"),
    booktabs = TRUE,
    escape = FALSE
  ) |>
  kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines")) |>
  kableExtra::column_spec(1, width = "1in") |>
  kableExtra::column_spec(2, width = "5in") |>
  kableExtra::pack_rows("\\agcurve{} curve parameters",
                        min(which(df_parameters$par_type == "ags")),
                        max(which(df_parameters$par_type == "ags")),
                        escape = FALSE) #|>
                        # kableExtra::pack_rows("x", 2, 2)

```