### 2024-04-09 ms/test.qmd
drafted tbl-models

---
title: Unknown
format:
  pdf:
    keep-tex: true  
    toc: false
header-includes: |
  \usepackage{hyperref}
  \addtokomafont{disposition}{\rmfamily}
  \usepackage{siunitx}
  \newcommand{\aax}{$\mathrm{AA}$}
  \newcommand{\Aamphi}{$A_{\mathrm{amphi}}$}
  \newcommand{\Ahypo}{$A_{\mathrm{hypo}}$}
  \newcommand{\agcurve}{$A \textendash g_\text{sw}$}
  \newcommand{\ca}{$C_\mathrm{a}$}
  \newcommand{\caequals}[1]{$C_\mathrm{a} = \qty{#1}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\cabetween}[2]{#1 < $C_\mathrm{a} < \qty{#2}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\gsw}{$g_\text{sw}$}
  \newcommand{\ppfd}{$\mathrm{PPFD}$}
  \newcommand{\ppfdequals}[1]{$\mathrm{PPFD} = \qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
  \newcommand{\rh}{$\mathrm{RH}$}
  \newcommand{\rhequals}[1]{$\mathrm{RH} = #1\%$}
  \newcommand{\tleafequals}[1]{$T_\mathrm{leaf} = \qty{#1}{\degreeCelsius}$}
mainfont: Times New Roman
bibliography: solanum-aa.bib
csl: https://www.zotero.org/styles/science
---

```{r load-objects, echo=FALSE, eval=TRUE}
library(dplyr)
library(readr)
library(tibble)
```

blah \autoref{tbl-models}.

```{r, models, echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-models
#| tbl-cap: "Summary of differences among competing models of how \\aax{} varies with light intensity, light treatment, and among accessions as a function of native \\ppfd. The models are numbered from simpler to more complex. All models include fixed effects of light intensity and light treatment; some models include interactions between. All models include a phylogenetic random effect of accession on \\aax; some models include varying effects of light intensity and light treatment among accessions. The last column indicates the accession-level \\aax{} variable we used as a response to native \\ppfd."

tribble(
    ~"Model", ~"Fixed effects", ~"Phylogenetic random effects", ~"Response to native \\ppfd",
    "1", "light intensity", "varying intercept among accessions", "$\\mathrm{AA}_{0, \\text{acc}}$",
    "1", "light treatment", "", "",
    "2", "light intensity", "varying intercept among accessions", "$\\mathrm{AA}_{0, \\text{acc}}$",
    "2", "light treatment", "", "",
    "2", "intensity $\\times$ treatment", "", "",
    "3", "light intensity", "varying intercept among accessions", "$\\mathrm{AA}_{2000, \\text{acc}}$",
    "3", "light treatment", "varying effect of high light intensity among accessions", "",
    "4", "light intensity", "varying intercept among accessions", "$\\mathrm{AA}_{\\text{sun, acc}}$",
    "4", "light treatment", "varying effect of sun treatment among accessions", "",
    "5", "light intensity", "varying intercept among accessions", "$\\mathrm{AA}_{2000, \\text{sun, acc}}$",
    "5", "light treatment", "varying effect of high light intensity among accessions", "",
    "5", "intensity $\\times$ treatment", "varying effect of sun treatment among accessions", ""
  ) |>
  knitr::kable(
    format = "latex",
    align = c("c", "l", "l", "l"),
    booktabs = TRUE,
    escape = FALSE,
    longtable = FALSE
  ) |>
    kableExtra::collapse_rows(columns = 1, latex_hline = "major", row_group_label_position = "first") |>
  kableExtra::column_spec(1, width = "0.3in") |>
  kableExtra::column_spec(2, width = "1.3in") |>
  kableExtra::column_spec(3, width = "3.4in") |>
  kableExtra::column_spec(4, width = "1in")

```