### 2025-10-19 ms/ms.qmd
updated ms with new analysis

---
title: Plasticity and adaptation to high light intensity amplify the advantage of amphistomatous leaves
format:
  pdf:
    keep-tex: true  
    toc: false
author:
  - name: "Christopher D. Muir$^{1,2*}$"
    orcid: 0000-0003-2555-3878
  - name: "Wei Shen Lim$^{1}$"
  - name: "Dachuan Wang$^{1,2}$"
affiliations: false
number-sections: false
header-includes: |
  \usepackage{hyperref}
  \addtokomafont{disposition}{\rmfamily}
  \usepackage{siunitx}
  \newcommand{\aahigh}{\mathrm{AA}_{2000}}
  \newcommand{\aalow}{\mathrm{AA}_{150}}
  \newcommand{\aashade}{\mathrm{AA}_{\text{shade}}}
  \newcommand{\aasun}{\mathrm{AA}_{\text{sun}}}
  \newcommand{\aax}{$\mathrm{AA}$}
  \newcommand{\aay}{\mathrm{AA}}
  \newcommand{\Aamphi}{$A_{\mathrm{amphi}}$}
  \newcommand{\Ahypo}{$A_{\mathrm{hypo}}$}
  \newcommand{\agcurve}{$A \textendash g_\text{sw}$}
  \newcommand{\aqcurve}{$A \textendash Q$}
  \newcommand{\ca}{$C_\mathrm{a}$}
  \newcommand{\caequals}[1]{$C_\mathrm{a} = \qty{#1}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\cabetween}[2]{#1 < $C_\mathrm{a} < \qty{#2}{\micro\mol\raiseto{-1}\mol}$}
  \newcommand{\gias}{$g_\text{ias}$}
  \newcommand{\gmax}{$g_\text{max}$}
  \newcommand{\gmaxratio}{$g_\text{max,ratio}$}
  \newcommand{\gmaxshade}{$g_\text{max,shade}$}
  \newcommand{\gmaxsun}{$g_\text{max,sun}$}
  \newcommand{\gsw}{$g_\text{sw}$}
  \newcommand{\loggsw}{$\log(g_\text{sw})$}
  \newcommand{\logA}{$\log(A)$}
  \newcommand{\ppfd}{$\mathrm{PPFD}$}
  \newcommand{\ppfdequals}[1]{$\mathrm{PPFD} = \qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
  \newcommand{\ppfdqty}[1]{$\qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
  \newcommand{\rh}{$\mathrm{RH}$}
  \newcommand{\rhequals}[1]{$\mathrm{RH} = #1\%$}
  \newcommand{\tleafequals}[1]{$T_\mathrm{leaf} = \qty{#1}{\degreeCelsius}$}
  \usepackage{setspace}
  \usepackage[modulo]{lineno}
mainfont: Times New Roman
bibliography: solanum-aa.bib
csl: https://www.zotero.org/styles/pnas
---

```{=latex}
\begin{center}
$^1$School of Life Sciences, University of Hawaiʻi; Mānoa, Hawaiʻi, USA\\
$^2$Department of Botany, University of Wisconsin, Madison; Wisconsin, USA\\
\medskip
*Corresponding author: cdmuir@wisc.edu\\
\medskip
ORCID: Christopher D. Muir — \href{https://orcid.org/0000-0003-2555-3878}{0000-0003-2555-3878}\\
\medskip
Classification: Biological Sciences, Ecology\\
\medskip
Keywords: amphistomy, photosynthesis, \textit{Solanum}, stomata
\end{center}
```

```{r load-objects, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(brms)
library(cowplot)
library(dplyr)
library(ggplot2)
library(glue)
library(kableExtra)
library(knitr)
library(lubridate)
library(magrittr)
library(readr)
library(stringr)
library(tibble)
library(tidyr)

theme_set(theme_cowplot())

source("../r/functions.R")

aa_args = read_rds("../objects/aa_args1.rds")
aa_stats = read_rds("../objects/aa_stats.rds")
# sort list by name
aa_stats = aa_stats[order(names(aa_stats))]

plant_info = read_rds("../data/plant-info.rds")

accession_info = read_rds("../data/accession-info.rds") |>
  dplyr::select(species, accession)

rh_curves = read_rds("../data/trimmed_rh_curves.rds")

stomata = read_rds("../data/stomata.rds")

fit_aa1 = read_rds("../objects/fit_aa1_sty.rds")
fit_aa2 = read_rds("../objects/fit_aa2_sty.rds")
dim_fit_aa = dim(fit_aa1$fit)

df_aa_pred1 = read_rds("../objects/df_aa_pred1.rds")
# df_aa_pred2 = read_rds("../objects/df_aa_pred2.rds")
df_aa_pred3 = read_rds("../objects/df_aa_pred3.rds")

aa_loo_table1 = read_rds("../objects/aa_sty_loo_table1.rds")
aa_loo_table2 = read_rds("../objects/aa_sty_loo_table2.rds")

n_plant = length(unique(rh_curves$acc_id))
n_acc = length(unique(rh_curves$acc))
n_sp = accession_info |>
  filter(accession %in% unique(rh_curves$acc)) |>
  pull(species) |>
  unique() |>
  length()

# average number of individual replicates for accession per light treatment
mean_n_rep = rh_curves |>
  summarize(
    n = length(unique(acc_id)),
    .by = c("acc", "light_treatment")
  ) |>
  pull(n) |>
  mean() |>
  prettyNum(digits = 3)

licor_date_range = plant_info |>
  filter(acc_id %in% unique(rh_curves$acc_id)) |>
  dplyr::select(`1s_rh_response_date`, `2s_rh_response_date`) |>
  pivot_longer(cols = everything()) |>
  pull(value) |>
  ymd() |>
  range(na.rm = TRUE)

n_weeks = prettyNum(as.numeric(difftime(licor_date_range[2], licor_date_range[1], units = "days")) / 7, digits = 3)
 # number of weeks for LICOR measurements
start_licor_date = format(licor_date_range[1], "%B %d, %Y")
end_licor_date = format(licor_date_range[2], "%B %d, %Y")

# Environmental data
df_germ_summary = read_rds("../data/df_germ_summary.rds") |>
  ungroup() |>
  summarize(value = median(value), .by = c("period", "variable", "parameter"))

df_growth_summary = read_rds("../data/df_growth_summary.rds") |>
  ungroup() |>
  filter(!is.na(value))

# environmental condition in germination room

germ_temp_day = df_germ_summary |>
  filter(period == "day", variable == "degC", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

germ_temp_night = df_germ_summary |>
  filter(period == "night", variable == "degC", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

germ_rh_day = df_germ_summary |>
  filter(period == "day", variable == "RH", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

germ_rh_night = df_germ_summary |>
  filter(period == "night", variable == "RH", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

germ_ppfd = df_germ_summary |>
  filter(period == "day", variable == "PAR", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

# environmental conditions in growth room

d1 = df_growth_summary |>
  summarize(value = median(value),
            .by = c("period", "rack", "variable", "parameter"))

ppfd_sun = d1 |>
  filter(period == "day", rack == "lower", variable == "PAR", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

ppfd_shade = d1 |>
  filter(period == "day", rack == "upper", variable == "PAR", parameter == "mu") |>
  pull(value) |>
  prettyNum(digits = 3)

# degrees warmer that the sun treatment air was
temp_diff = df_growth_summary |>
  filter(period == "day", rack != "hobo", variable == "degC", parameter == "mu") |>
  pivot_wider(names_from = "rack") |>
  mutate(diff = lower - upper) |>
  summarize(value = median(diff, na.rm = TRUE)) |>
  pull(value) |>
  prettyNum(digits = 3)

# difference in RH
rh_diff = df_growth_summary |>
  filter(period == "day", rack != "hobo", variable == "RH", parameter == "mu") |>
  pivot_wider(names_from = "rack") |>
  mutate(diff = upper - lower) |>
  summarize(value = median(diff, na.rm = TRUE)) |>
  pull(value) |>
  prettyNum(digits = 3)

df_leaf_temp = plant_info |>
  filter(!is.na(leaf_temperature_C)) |>
  summarize(
    leaf_temperature_C = mean(leaf_temperature_C),
    n = n(),
    .by = "light_treatment"
  )

# difference in leaf temperature between treatments
leaf_temp_diff = df_leaf_temp$leaf_temperature_C |>
  diff() |>
  prettyNum(digits = 3)

# number of leaves for which we have measured leaf temp under ambient conditions
n_leaf_temp = sum(df_leaf_temp$n)

# data on leaflet type
df_leaflet_type = plant_info |>
  summarize(n = n(), .by = "leaflet") |>
  pivot_wider(names_from = "leaflet", values_from = "n") |>
  mutate(
    percent_terminal = prettyNum(terminal / (terminal + lateral) * 100, digits = 3),
    percent_lateral = prettyNum(lateral / (terminal + lateral) * 100, digits = 3)
  )

percent_terminal = df_leaflet_type$percent_terminal
percent_lateral = df_leaflet_type$percent_lateral
# percent of time that anatomical traits were measured on different leaflet than LICOR measurements
percent_change_leaflet = full_join(
  plant_info |>
    filter(acc_id %in% unique(rh_curves$acc_id)) |>
    select(acc_id, licor = leaflet),
  stomata |>
    mutate(acc_id = str_c(accession, replicate, sep = "-")) |>
    filter(acc_id %in% unique(rh_curves$acc_id)) |>
    select(acc_id, stomata = leaflet),
  by = join_by(acc_id)
) |>
  summarize(x = mean(licor != stomata, na.rm = TRUE)) |>
  pull(x) |>
  multiply_by(100) |>
  prettyNum(digits = 3)

 # percent of times where AA and AQ curves were done on same leaflet
percent_aa_and_aq = plant_info |>
  filter(acc_id %in% unique(rh_curves$acc_id)) |>
  mutate(same_leaflets = !switch_leaflets) |>
  pull(same_leaflets) |>
  mean() |>
  multiply_by(100) |>
  prettyNum(digits = 3)

# data on chamber light conditions
percent_red = rh_curves |>
  mutate(p_red = Q_red / Q) |>
  summarize(p_red = mean(p_red, na.rm = TRUE),
            .by = c(acc_id, light_intensity)) |>
  summarize(p_red = mean(p_red, na.rm = TRUE), .by = light_intensity) |>
  pivot_wider(names_from = light_intensity, values_from = p_red)

p_red_high = format(signif(100 * percent_red$`2000`, 3), nsmall = 1)
p_blue_high = format(signif(100 * (1 - percent_red$`2000`), 3), nsmall = 1)
p_red_low = format(signif(100 * percent_red$`150`, 3), nsmall = 1)
p_blue_low = format(signif(100 * (1 - percent_red$`150`), 3), nsmall = 1)

# percent measure LMA on alternate leaflet
percent_lma_switch = plant_info |>
  filter(acc_id %in% rh_curves$acc_id) |> 
  summarize(x = mean(leaflet != lma_leaflet, na.rm = TRUE)) |>
  pull(x) |>
  multiply_by(100) |>
  prettyNum(digits = 3)

 # percent leaves missing LMA
percent_lma_missing = plant_info |>
  filter(acc_id %in% rh_curves$acc_id) |> 
  summarize(x = mean(is.na(lma_leaflet))) |>
  pull(x) |>
  multiply_by(100) |>
  prettyNum(digits = 3)

### Cleaning RH curves
ca_min = aa_args$ca_min
ca_max = aa_args$ca_max
n_min_points_per_rh_curve = aa_args$n_min_points_per_rh_curve
point_outlier_threshold = aa_args$point_outlier_threshold
aa_outlier_threshold = aa_args$aa_outlier_threshold
thinning_interval = aa_args$thinning_interval

### Cleaning AQ curves
aq_r2_threshold = aa_args$aq_r2_threshold

### Bayesian data analysis
n_model1 = nrow(aa_loo_table1)
n_model2 = nrow(aa_loo_table2)

### Summarizing AA results
aa_mean_range = df_aa_pred1 |>
  ungroup() |>
  mutate(signif = sign(.lower) == sign(.upper)) |>
  summarize(
    min = min(aa),
    max = max(aa),
    aa = mean(aa),
    n_signif = sum(signif),
    n_acc = n(),
    .by = c(light_treatment, light_intensity)
  ) |>
  mutate(x = sprintf(
    "%.3f (range: %.3f–%.3f; %.0f of %.0f populations significant)",
    aa,
    min,
    max,
    n_signif,
    n_acc
  ))

aa_summary3 = df_aa_pred3 |>
  mutate(signif = (.upper < 0)) |>
  summarize(n_signif = sum(signif), .by = light_treatment) |>
  pull(n_signif, light_treatment)

# Effect of light treatment on LMA
post_lma = posterior_summary(fit_aa2)["b_llma_light_treatmenthigh", ]

## PAI results
df_coef_summary = read_rds("../objects/df_coef_summary.rds") |>
  filter(term == "slope") |>
  mutate(signif = (sign(.lower) == sign(.upper))) |>
  summarize(n_signif= sum(signif),
            n_total = n()) |>
  as.list()

```

```{=latex}
\newpage
\begin{linenumbers}
\begin{doublespace}
```

## Abstract

<!-- Abstracts of Research Articles should explain to the general reader why the research was done, what was found and why the results are important. They should start with some brief BACKGROUND information: a sentence giving a broad introduction to the field comprehensible to the general reader, and then a sentence of more detailed background specific to your study. This should be followed by an explanation of the OBJECTIVES/METHODS and then the RESULTS. The final sentence should outline the main CONCLUSIONS of the study, in terms that will be comprehensible to all readers. The Abstract is distinct from the main body of the text and thus should not be the only source of background information critical to understanding the manuscript. Please do not include citations or abbreviations in the Abstract. The abstract should be 125 words or less. For Perspectives and Policy Forums, please include a one-sentence abstract. -->

The presence of stomata on both leaf surfaces (amphistomy) increases photosynthesis by reducing the distance for CO$_2$ diffusion between stomata and chloroplasts. Paradoxically, most leaves are hypostomatous (stomata on lower surface), despite the photosynthetic advantage of amphistomy. Across `r n_acc` diverse populations of "wild tomatoes", leaves developed under high light intensity benefit more from amphistomy in terms of CO$_2$ assimilated for a given stomatal conductance than plants developed under low light. Furthermore, populations native to open habitats benefit more from amphistomy than those from more closed habitats. Thus, plasticity and adaptation together may explain why amphistomatous leaves are prevalent in sunny, open habitats, including many crops. Contrary to common assumptions, amphistomy can save water because hypostomatous leaves evaporate 10-65\% more to achieve the same photosynthetic rate.

 <!-- LONG FORM - save in case of resubmission -->
<!-- The presence of stomata on both leaf surfaces (amphistomy) can increase photosynthesis in C$_3$ plants by reducing the path length for CO$_2$ diffusion between substomatal cavities and chloroplasts. Paradoxically, despite this photosynthetic advantage, most leaves are hypostomatous (stomata on lower surface only). Based on the observation that amphistomatous leaves are most common among herbaceous plants growing in sunny habitats, including many crop species, it has long been hypothesized that traits affecting CO$_2$ supply or demand amplify the advantage of amphistomy under high light intensity. We used a recently developed method to quantify amphistomy advantage, the photosynthetic rate of an amphistomatous leaf relative to an otherwise identical hypostomatous leaf, in `r n_acc` diverse populations representing `r n_sp` species of wild tomatoes (*Solanum* sect. *Lycopersicon* and sect. *Lycopersicoides*). Plants grown under high light intensity benefit more from amphistomy in terms of CO$_2$ assimilated for a given stomatal conductance than plants grown under low light, regardless of measurement light intensity. Furthermore, populations native to open habitats tend to benefit more from amphistomy compared to those from more closed habitats. Developmental plasticity and adaptation of leaves to high light intensity likely play a combined role in explaining the adaptive significance and distribution of amphistomy in nature. We estimate that hypostomatous leaves would expend 10-65\% more water to achieve the same photosynthetic rate as a comparable amphistomatous leaf, implying large fitness costs of amphistomy in ecological contexts where this trait is rare. -->

## Significance statement

Stomata, the "mouths" of plants, regulate CO$_2$ available for photosynthesis. They are ubiquitous on leaves and essential to understanding why vascular plants transformed Earth and how our crops grow. Yet the adaptive significance of why many leaves only have stomata on their lower surface is a mystery. Many assume that this limits water loss, but we show the opposite is true. By manipulating which stomata a leaf can use, we show that leaves with stomata on both surfaces gain more CO$_2$ for the same amount of water lost, increasing water use efficiency. Furthermore, species from open canopies grown under high light benefit the most, which may explain why leaves with stomata on both sides are most prevalent in these habitats.

## Main text

<!-- Paragraph 1 -->
Stomata are microscopic pores on the surfaces of leaves and other photosynthetic organs formed by a pair of guard cells. They are essential for balancing carbon gained per unit water lost and were a key innovation in vascular plants, permitting them to grow tall on land by enabling access to CO$_2$ for photosynthetic carbon assimilation while preventing hydraulic failure in variable environments [@raven_selection_2002; @mcadam_stomata_2021; @clark_origin_2022]. Optimal stomatal function depends on both dynamic changes in aperture on the scale of minutes to hours, as well as static anatomy determined by developmental plasticity and constitutive genetic differences [@hetherington_role_2003; @de_boer_optimal_2016; @harrison_influence_2020]. Understanding how stomata respond to environmental change over daily, developmental, and evolutionary time is important for studying adaptation [@buckley_modelling_2013; @raven_selection_2002; @haworth_co-ordination_2013], inferring paleoclimate from fossil cuticles [@franks_new_2014; @mcelwain_paleoecology_2017; @the_cenozoic_co_proxy_integration_project_cencopip_consortium_toward_2023], predicting responses to climate change [@woodward_stomatal_1987; @liang_stomatal_2023; @chua_stomatal_2024; @lang_century-long_2024], and improving crops [@hofmann_impact_2025]. Stomatal function contributes to global carbon and water cycles [@berry_stomata:_2010] and therefore to predicting future climate [@franks_stomatal_2017].

<!-- Paragraph 2 -->
<!-- Parkhurst gradients 1988 -->
<!-- Syvertsen et al 1995 - may be more here to check about gi in sun and shade-->
<!-- "Farquhar and Raschke (2). However, using a technique similar to ours, Sharkey et al. (11) estimate r, in Xan:hium to be closer to 1.0 m2 s mol-', and state that they believe the estimate of 3.2 m2 s mol-' to be high" -->
Despite extensive theoretical and empirical progress understanding stomatal function and anatomy from molecular to ecosystem levels, the adaptive significance of amphistomatous leaves remains an important unsolved problem in leaf structure-function relationships [@grubb_leaf_1977; @parkhurst_adaptive_1978; @mott_adaptive_1982; @gibson_structure-function_1996; @smith_leaf_1997; @oguchi_leaf_2018; @drake_two_2019; @grubb_leaf_2020]. Amphistomatous leaves develop abaxial and adaxial stomata whose aperture can be independently regulated [@pospisilova_environmental_1980; @mott_stomatal_1984; @reich_effects_1985; @mott_asymmetric_1993; @wall_stomata_2022] to control gas exchange through each surface. All else being equal, simultaneous gas exchange through stomata on both surfaces increases CO$_2$ supply to chloroplasts by providing a second parallel pathway through leaf intercellular airspaces, enhancing photosynthesis [@parkhurst_adaptive_1978; @gutschick_photosynthesis_1984]. The extent to which amphistomy increases CO$_2$ supply depends on resistance to diffusion in intercellular airspaces. This resistance can be low in thin, porous, amphistomatous leaves [@mott_stomatal_1984; @marquez_assessing_2023], but may be more substantial in thick, dense, hypostomatous leaves [@parkhurst_intercellular_1990]. We refer to the intercellular airspace conductance (\gias), the inverse of resistance. Amphistomatous leaves also lose more water through evaporation because of a second boundary layer conductance [@foster_influence_1986], but the additional carbon gain should be enough to offset this cost in most realistic scenarios [@muir_is_2019].

<!-- Paragraph 3 -->
The paradoxical fact is that, despite the photosynthetic benefit, most leaves are not amphistomatous. Many vertically oriented and/or isobilateral leaves are amphistomatous [@wood_physiology_1934; @howell_concerning_1945, @drake_two_2019]. But among dorsiventral leaves, it is primarily herbaceous plants in open, high light habitats that tend to have amphistomatous leaves [@salisbury_i_1928; @peat_comparative_1994; @gibson_structure-function_1996; @jordan_using_2014; @bucher_stomatal_2017; @muir_light_2018; @triplett_stomatal_2025]. Most other leaves, except those from aquatic habitats, are hypostomatous, producing stomata only on the lower, abaxial surface. Even resupinate leaves develop stomata on the lower, albeit adaxial surface [@lyshede_comparative_2002], suggesting that leaf orientation (lower vs. upper) rather than leaf polarity (abaxial vs. adaxial) is causal. Stomatal density ratio, defined as the ratio of upper to total stomatal density, is a quantitative metric of stomatal distribution. The covariation between stomatal density ratio and light habitat is both qualitative and quantitative. A higher proportion of sun leaves are amphistomatous [@salisbury_i_1928] and the proportion of stomata on the upper, adaxial surface increases with light [@bucher_stomatal_2017; @muir_light_2018]. Resolving why high light intensity favors amphistomatous dorsivental leaves is an important first step toward understanding variation in stomatal density ratio and leaf structure-function relationships more generally.

<!-- Paragraph 4+, hypotheses and predictions -->
The overarching hypothesis is that leaves with greater stomatal density ratio are more common in open, sunny habitats because they increase photosynthesis most in those circumstances. An amphistomatous leaf increases photosynthetic carbon gain compared to an otherwise identical hypostomatous leaf by increasing conductance through the leaf intercellular airspaces and boundary layers. We quantify this benefit as the amphistomy advantage (\aax{}), the log-response ratio of photosynthesis in an amphistomatous leaf compared to an otherwise identical pseudohypostomatous leaf [@parkhurst_adaptive_1978; @triplett_amphistomy_2024]. Why would \aax{} be greater in sun than shade? We consider three nonmutually exclusive hypotheses that we classify as 'acclimation', 'plasticity', and 'constitutive' ([Fig. @fig-hypotheses]).

*Acclimation hypothesis*: Photosynthetic induction to high light intensity typically involves increases in total leaf stomatal conductance (increased CO$_2$ supply), the concentration of active Rubisco, and electron transport capacity (increased CO$_2$ demand). If the acclimation hypothesis is correct, we predict that $\aahigh{} > \aalow{}$ for all populations regardless of native habitat or growth environment. $\aahigh{}$ is the \aax{} measured under high light intensity (\ppfdequals{2000}); $\aalow{}$ is the \aax{} measured under low light intensity (\ppfdequals{150}).

*Plasticity hypothesis*: Individuals of the same genotype often develop dramatically different leaves in sun and shade conditions [@poorter_metaanalysis_2019]. Plastic responses are likely adaptations to optimize photosynthesis at different light intensities in variable environments [@givnish_common-garden_2014]. Plastic changes in leaf anatomy and biochemistry could modulate \aax{} as a byproduct. Thicker or less porous leaves, both of which are associated with high leaf mass per area (LMA), will have lower \gias; leaves with increased total stomatal density and photosynthetic capacity have greater potential CO$_2$ supply and demand. Under the plasticity hypothesis, we predict that $\aasun{} > \aashade{}$ for all populations and light intensities. $\aasun{}$ is the \aax{} measured on sun leaves; $\aashade{}$ is the \aax{} measured on shade leaves.

*Constitutive hypothesis*: In environments that are relatively constant or where environmental change cannot be anticipated by a reliable cue, natural selection will favor constitutive expression of optimal phenotypes. We therefore predict genotypes from more sunny, open habitats will have consistently greater \aax{} under all measurement and growth light intensities. For herbaceous plants, light intensity is largely a function of the tree canopy [@engelbrecht_evaluation_2001]. Herbs growing in the open will regularly experience high light intensity; herbs growing under a forest canopy will often experience low light intensity.

```{r, hypotheses, echo = FALSE}
#| label: fig-hypotheses
#| fig-align: center
#| fig-cap: "\\textbf{Conceptual outline of three nonmutually exclusive hypotheses explaining why amphistomy advantage (\\aax{}) might be greater for leaves in sunny, open habitats.} The acclimation hypothesis predicts that \\aax{} is greater under high measurement light intensity (\\ppfdequals{2000}) than low measurement light intensity (\\ppfdequals{150}), regardless of growth light intensity or native plant area index (PAI). The plasticity hypothesis predicts that \\aax{} is greater for plants grown in sun than shade, regardless of measurement light intensity or native PAI. The constitutive hypothesis predicts that \\aax{} is greater for plants adapted to sunny habitats (lower native PAI) than shaded habitats, regardless of measurement light intensity or growth light intensity."

knitr::include_graphics("../figures/hypotheses.pdf")
```

The primary directional predictions for each hypothesis are summarized in @tbl-hypotheses; detailed predictions for results that would indicate simultaneous support for multiple hypotheses are in @tbl-predictions.

```{r, hypotheses, echo = FALSE}
#| label: tbl-hypotheses
#| tbl-cap: "Three nonmutually exclusive hypotheses and directional predictions explaining why amphistomy advantage (\\aax{}) might be greater for leaves in sunny, open habitats. For each hypothesis, we make predictions for how measurement light intensity (\\ppfdequals{150} vs. \\ppfdqty{2000}), growth light intensity (sun vs. shade), and native plant area index (PAI) would affect \\aax{}. \\ppfd{}: photosynthetic photon flux density."

kableExtra::kable(
  data.frame(
    hypothesis = c("acclimation", "plastic", "constitutive"),
    measurement = c(
      "$\\mathrm{AA}_{2000} > \\mathrm{AA}_{150}$",
      "$\\mathrm{AA}_{2000} = \\mathrm{AA}_{150}$",
      "$\\mathrm{AA}_{2000} = \\mathrm{AA}_{150}$"
    ),
    growth = c(
      "$\\mathrm{AA}_{\\text{sun}} = \\mathrm{AA}_{\\text{shade}}$",
      "$\\mathrm{AA}_{\\text{sun}} > \\mathrm{AA}_{\\text{shade}}$",
      "$\\mathrm{AA}_{\\text{sun}} = \\mathrm{AA}_{\\text{shade}}$"
    ),
    native = c(
      "$\\operatorname{cor}(\\mathrm{PAI}, \\mathrm{AA}) = 0$",
      "$\\operatorname{cor}(\\mathrm{PAI}, \\mathrm{AA}) = 0$",
      "$\\operatorname{cor}(\\mathrm{PAI}, \\mathrm{AA}) < 0$"
    )
  ),
  col.names = c(
    "Hypothesis",
    "Measurement light intensity",
    "Growth light intensity",
    "Native PAI"
  ),
  escape = FALSE,
  booktabs = TRUE,
  format = "latex",
  align = "c"
) |>
  kable_styling() |>
  column_spec(1, width = "1in") |>
  column_spec(2, width = "1.5in") |>
  column_spec(3, width = "1.5in") |>
  column_spec(4, width = "1.25in")
```

<!-- Paragraph 5 (summarize methods) -->
We tested these hypotheses by comparing \aax{} among amphistomatous wild tomato species (@tbl-populations; @fig-sr-rxnorm; [@peralta_taxonomy_2008]) from different native light habitats, grown under simulated sun and shade light treatments, and measured under contrasting light intensities (low and high). We measured \aax{} on `r n_plant` individual plants from `r n_acc` populations (average of `r mean_n_rep` replicates per light treatment) using a recently developed method [@triplett_amphistomy_2024]. With this method, we directly compare the photosynthetic rate of an untreated amphistomatous leaf to that of the same leaf with gas exchange blocked through the adaxial (upper) surface by transparent plastic, which we refer to as 'pseudohypostomy'. To compare amphi- and pseudohypostomatous leaves at identical whole-leaf stomatal conductance (\gsw), we measure $A$ over a range of \gsw, inducing stomatal opening and closure by modulating humidity (see supplementary text for further details). We estimated 'amphistomy advantage' (\aax) *sensu* @parkhurst_adaptive_1978, but with modifications previously described in @triplett_amphistomy_2024 and here (Supplementary Materials). The native light intensity was represented by plant area index (PAI $\unit{\meter\squared\per\meter\squared}$), estimated using a global gridded data set derived from the Global Ecosystem Dynamics Investigation [GEDI; @burns_multi-resolution_2024] and georeferenced accession collection information from the Tomato Genetics Resource Center (@tbl-populations). The growth light intensities were \ppfdequals{`r ppfd_sun`} (sun treatment) and \ppfdqty{`r ppfd_shade`} (shade treatment) while all other environment conditions were nearly identical (see Materials and Methods). The high and low measurement light intensities were \ppfdequals{2000} (`r p_red_high`:`r p_blue_high` red:blue) and \ppfdequals{150} (`r p_red_low`:`r p_blue_low` red:blue), respectively.

<!-- Paragraph 6 (summarize main results) -->

```{r, echo = FALSE}
x = if (df_coef_summary[['n_signif']] < df_coef_summary[['n_total']]) {
  glue("{n1} of the {n2}", n1 = df_coef_summary[['n_signif']], n2 = df_coef_summary[['n_total']])
} else {
  "all"
}
```

Consistent with biophysical theory of CO$_2$ diffusion within leaves, $\aay{} > 0$ for all populations ([Fig. @fig-aa]A). We tested whether \aax{} varied among treatments by comparing the fit of competing Bayesian phylogenetic mixed effects models using information criteria (see Materials and Methods). Models including measurement and growth light intensity consistently outperformed simpler models (@tbl-aa_loo1). There was equivocal support for interactions between factors and population because several models including these terms performed similarly to those without them (@tbl-aa_loo1). We focus on results that were robust to model structure and addressed our *a priori* hypotheses using parameter estimates from the best model.

Measured under high light intensity, \aax{} was consistently greater for sun plants. The average \aax{} among populations in the shade treatment was `r filter(aa_mean_range, light_treatment == "low", light_intensity == "2000")$x`; however, the same populations grown at high light intensity showed a mean \aax{} of `r filter(aa_mean_range, light_treatment == "high", light_intensity == "2000")$x`. Contrary to the predictions of the acclimation hypothesis, \aax{} was greater in all populations under low measurement light intensity for both sun and shade grown plants. The overall average \aax{} of shade and sun grown plants measured under low light intensity was `r filter(aa_mean_range, light_treatment == "low", light_intensity == "150")$x` and `r filter(aa_mean_range, light_treatment == "high", light_intensity == "150")$x`, respectively. There was a modest tendency for populations from more open habitats (lower PAI) to exhibit greater \aax{} and the slope was significantly different than 0 in `r x` treatment combinations ([Fig. @fig-aa]B).

```{r, aa, echo = FALSE, warning=FALSE, message=FALSE} 
#| label: fig-aa
#| fig-align: center
#| fig-width: 6
#| fig-height: 8
#| fig-cap: "Caption on following page."

plot_grid(
  read_rds("../objects/fig_aa.rds"),
  read_rds("../objects/fig_aa_pai.rds"),
  ncol = 1,
  labels = "AUTO",
  label_size = 20,
  label_fontface = "bold",
  align = "hv"
)

```

```{=latex}
\addtocounter{figure}{-1}
```

```{r, aa_cap, echo = FALSE, warning=FALSE, message=FALSE} 
#| label: fig-aacap
#| fig-align: center
#| fig-width: 0.1
#| fig-height: 0.1
#| fig-cap: "\\textbf{Amphistomy advantage (\\aax{}) is greater in wild tomatoes from open habitats and is further amplified by developmental plasticity under sunny growth conditions.} \\aax{} ($y$-axis) is the log-response ratio of photosynthesis in an amphistomatous leaf compared to an otherwise identical pseudohypostomatous leaf. In both panels, estimates are shown for plants grown under simulated shade (brown) or sun (orange) and measured under high light intensity (\\ppfdequals{2000}; open circles) and low light intensity (\\ppfdequals{150}; solid circles). The dashed line at zero indicates no difference in photosynthesis between amphi- and pseudohypostomatous leaves. (\\textbf{A}) The points are the posterior median \\aax{} for each population, with error bars showing 95% confidence intervals. Within each facet, the populations are arranged by \\aax{} estimated in that growth and measurement condition. The average \\aax{} among populations for each treatment group is written in the top left corner. (\\textbf{B}) The same estimates of \\aax{} in each combination of growth and measurement light intensity plotted against native plant area index (PAI; $x$-axis, log-scale). The confidence intervals were omitted for visual clarity. The median relationship between native PAI ($x$-axis; log-scale) and \\aax{} is shown as a solid line with the shaded region between dashed lines showing the 95% confidence ribbon. The * in the top left corner of each facet indicates the slope of the linear regression is significantly different from zero."

# Add to caption if needed
# "; n.s. indicates not significant."

ggplot()
```

<!-- Paragraph 7 (summarize secondary results) -->

<!-- Text from previous version where effect of growth light intensity was retained in LMA model:

While LMA is weakly, albeit significantly (@tbl-fit_aa2), associated with individual-level \aax{} ([Fig. @fig-lma_aa]), the effect of growth light intensity on \aax{} is still predictive based on model comparison using information criteria (@tbl-aa_loo1). However, the direct effect of the sun treatment on \aax{} was weaker when LMA was included in the model (@tbl-fit_aa1 vs. @tbl-fit_aa2), suggestive of a mediating role.

-->

The pattern of \aax{} across wild tomatoes strongly supports the plasticity hypothesis, contradicts the acclimation hypothesis, and provides modest support for the constitutive hypothesis. Plastic changes in leaf thickness and/or packing density, summarized by the bulk leaf mass per area (LMA), may mediate the effect of growth light intensity on \aax{}. LMA increased in sun grown plants in all populations by an average of `r signif(100 * (exp(post_lma["Estimate"]) - 1), 3)`% [95% CI: `r signif(100 * (exp(post_lma["Q2.5"]) - 1), 3)` to `r signif(100 * (exp(post_lma["Q97.5"]) - 1), 3)`%], qualitatively similar to plastic responses in many species [@poorter_metaanalysis_2019]. When LMA is included in statistical models, the effect growth light intensity on \aax{} is no longer supported, indicating that plastic changes in LMA mediate the effect of growth light intensity (@tbl-fit_aa2). The effect of LMA on individual-level \aax{} ([Fig. @fig-lma_aa]) is weak, albeit significantly (@tbl-fit_aa2). This is likely because many anatomical traits underlie LMA [@john_anatomical_2017] and future research will be needed to identify which particular traits, such as leaf thickness or mesophyll porosity, are responsible for mediating \aax{}. The fact that the \aax{} of sun plants was greater under low measurement light intensity supports a long-standing hypothesis that resistance to CO$_2$ diffusion is greater in the upper than lower portions of the leaf interior [@jones_effects_1972]. At low light intensity, photosynthesis is weighted toward the upper palisade where most light is intercepted. If resistance to CO$_2$ diffusion is high in the upper palisade, amphistomy may, unexpectedly, be particularly beneficial for sun leaves experiencing intermittent shade or cloud cover. Our study is limited in testing this because we could not directly measure the stomatal conductance ratio and intercellular resistance on each surface. Future experiments measuring \aax{} with a dual sided chamber [@wall_stomata_2022; @marquez_assessing_2023] can overcome these limitations.

```{r, lma_aa, echo = FALSE,warning=FALSE, message=FALSE} 
#| label: fig-lma_aa
#| fig-align: center
#| fig-width: 6
#| fig-height: 4
#| fig-cap: "\\textbf{Plastic changes in leaf mass per area (LMA) partially mediate the effect of sun- to shade-grown plants on amphistomy advantage (\\aax{}).} Each point is the estimated LMA ($x$-axis) and \\aax{} ($y$-axis) for each population grown under high light intensity (\\ppfdequals{2000}) and low light intensity (\\ppfdequals{150}). The points are colored by growth light treatment. The solid line is the posterior median of the linear regression of individual-level LMA on $\\aahigh$  with the shaded region showing the 95% confidence ribbon; the dashed line is the same for $\\aalow$. Confidence intervals for each population estimate were omitted for visual clarity."

read_rds("../objects/fig_lma_aa.rds")

```

We conclude that both developmental plasticity and adaptation to open habitats, but not acclimation, may explain the long-standing observation that amphistomatous leaves are more common in sunny habitats, at least among herbaceous plants including crop relatives. This result changes our understanding by showing that high light intensity *per se* does not increase the benefit of amphistomy. Instead, the benefit increases as a byproduct of anatomical and biochemical changes caused by plasticity and adaptation to higher light intensity. To gain more precise understanding of when amphistomy is most beneficial will require further research on its leaf anatomical and biochemical basis.

The magnitude of \aax{} we document across populations is also noteworthy, as it implies that hypostomatous leaves predominant in mesic to wet forests globally are giving up 'free' carbon that would require little to no additional water loss. The cost of hypostomy can be quantified as the difference in total \gsw{} that an amphistomatous leaf would require to achieve the same photosynthetic rate. This cost can be locally approximated on a log-ratio scale as $\mathrm{AA} / \varepsilon_g$ (Supporting Information), where $\varepsilon_g$ is the elasticity of $A$ to \gsw. For $\mathrm{AA} = 0.05$, the water cost of hypostomy would be 0.5 (64.9\%) at low elasticities ($\varepsilon_g = 0.1$) and 0.1 (10.5\%) at high elasticities ($\varepsilon_g = 0.5$). For plants to expend this much extra water implies a large fitness cost of upper stomata in certain ecological contexts. As this is the first quantitative comparison of \aax{} across species, estimates from a broader range of species and more stressful environments will be required to fully understand the ecological causes and consequences of stomatal distributions on leaf surfaces.

## Acknowledgements

Sam McKlin and Tom Buckley helped with protocol development. Justin Alter, Max Gatlin, Joana Kim, Jenna Matsuyama, Brandon Najarian, and Kai Yasuda contributed to data collection. Kate McCulloh provided feedback on an earlier version of this manuscript. Sarah Friedrich helped with graphics. We used Copilot and ChatGPT to assist with coding.

### Funding:

US National Science Foundation OIA-1929167 (C.D.M.)

### Author contributions:

Conceptualization: C.D.M.; Methodology: C.D.M., W.S.L.; Investigation: C.D.M., W.S.L., D.W.; Visualization: C.D.M.; Funding acquisition: C.D.M.; Writing – original draft: C.D.M.; Writing – review & editing: C.D.M., W.S.L., D.W.

```{=latex}
\end{doublespace}
\end{linenumbers}
```

## References

::: {#refs}
:::

{{< pagebreak >}}

# Supplementary Materials

```{=latex}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\theequation}{S\arabic{equation}}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{equation}{0}
\setcounter{page}{1}
```

## Materials and Methods {#sec-methods}

### Populations

We compared \aax{} among `r n_acc` ecologically diverse populations of wild tomato, including representatives of all described species of *Solanum* sect. *Lycopersicon* and sect. *Lycopersicoides* [@peralta_taxonomy_2008] and the cultivated tomato *S. lycopersicum* var. *lycopersicum* (@tbl-populations). These populations were selected in a previous study because they encompass the breadth of climatic variation in the wild tomato clade [@pease_phylogenomics_2016]. In one case, we substituted a congeric population of the species (*S. galapagense*) because of difficult growing the focal population. We also added an a population of *S. pennellii*. Due to constraints on growth space and time, we spread out measurements over `r n_weeks` weeks from `r start_licor_date` to `r end_licor_date`. Replicates within population were evenly spread out over this period to prevent confounding of temporal variation in growth conditions with variation among populations.

```{r, populations, eval=TRUE,echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-populations
#| tbl-cap: "Accession information of *Solanum* populations used in this study. The species name, accession number, collection latitude, longitude, elevation, and plant area index (PAI) from the Global Ecosystem Dynamics Investigation (see 'Climate data'). TGRC: Tomato Genetics Resource Center."

# Copied from solanum-aq project

accession_info = read_rds("../data/accession-info.rds") |>
  filter(accession %in% rh_curves$acc) |>
  left_join(read_rds("../data/accession-gedi.rds") |>
              select(accession, pai), by = join_by(accession)) |>
  arrange(species, accession) |>
  select(
    Species = species,
    `TGRC accession` = accession,
    `Latitude` = lat,
    `Longitude` = lon,
    `Elevation (mas)` = elev,
    `$\\mathrm{PAI}~(\\unit{\\meter\\squared\\per\\meter\\squared})$` = pai
  )

accession_info |>
  knitr::kable(
    format = "latex",
    digits = c(rep(0, 2), 3, 3, 0, 2),
    booktabs = TRUE,
    longtable = TRUE,
    escape = FALSE
  ) |>
  kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines", "repeat_header")) |>
  kableExtra::column_spec(1, width = "3.25cm", italic = TRUE)

```

### Plant growth conditions

We recorded $\mathrm{PPFD}$ using full spectrum quantum sensors (SQ-500-SS, Apogee Instruments, Logan, Utah, USA); we recorded temperature, RH, and [CO$_2$] using an EE894 sensor (E+E Elektronik, Engerwitzdorf, Austria) protected by a radiation shield. All environmental measurements were taken every 10 minutes from the middle of plant racks at approximately the same height as the leaves we measured. We measured leaf temperature of focal leaves prior to measurement using an infrared radiometer (SI-111-SS, Apogee Instruments, Logan, Utah, USA).

#### Germination and seedling stage

Seeds provided by the Tomato Genetics Resource Center germinated on moist paper in plastic boxes after soaking for 30-60 minutes in a 50% (volume per volume) solution of household bleach and water, followed by a thorough rinse. We transferred seedlings to cell-pack flats containing Pro-Mix BX potting mix (Premier Tech, Rivière-du-Loup, Quebec, Canada) once cotyledons fully emerged, typically within 1-2 weeks of sowing. We grew seeds and seedlings for both sun and shade treatments under the same environmental conditions (12:12 h, `r germ_temp_day`:$\qty{`r germ_temp_night`}{\degreeCelsius}$, `r germ_rh_day`:`r germ_rh_night` RH day:night cycle). LED light provided $\mathrm{PPFD} = \qty{`r germ_ppfd`}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$ (Fluence RAZRx, Austin, Texas, USA) during the germination and seedling stages.

#### Light treatments

Seedlings were randomly assigned in alternating order within population to the sun or shade treatment during transplanting. After seedlings established in cell-pack flats for $\approx 2$ weeks, we transplanted them to 3.78 L plastic pots containing 60% Pro-Mix BX potting mix, 20% coral sand (Pro-Pak, Honolulu, Hawaiʻi, USA), and 20% cinders (Niu Nursery, Honolulu, Hawaiʻi, USA). Percentage composition is on a volume basis. The soil mixture contained slow release NPK fertilizer following manufacturer instructions (Osmocote Smart-Release Plant Food Flower & Vegetable, The Scotts Company, Marysville, Ohio, USA). We determined pot field capacity one week after transplanting using a scale (Ohaus V12P15 Valor 1000, Parsippany, New Jersey, USA) and watered to field capacity three times per week to prevent drought stress.

We assigned sun and shade treatment to lower and upper racks of a $\qty{1.22}{\meter} \times \qty{2.44}{\meter}$ shelving unit in a climate-controlled growth room. We assigned the sun treatment to the lower rack to limit diffuse light from reaching the shade treatment. The average daytime \ppfd{} was \ppfdqty{`r ppfd_sun`} and \ppfdqty{`r ppfd_shade`} for sun and shade treatments, respectively. To isolate the effect of light intensity from quality, we used the same LED model with the the same spectrum (Fluence SPYDR 2i, Austin, Texas, USAS), but dimmed the lights in the shade treatment. To maintain homogeneous environmental conditions other than light, we mixed air within the growth room using an air circulator (Vornado 693DC, Andover, Kansas, USA) and within racks using a miniature oscillating air circulator (Vornado Atom 1, Andover, Kansas, USA). Despite these efforts, the air in the sun treatment was on average $\qty{`r temp_diff`}{\degreeCelsius}$ warmer and the average RH was consequently `r rh_diff`% lower. However, because of evaporative cooling, the leaves in the sun treatment were only $\qty{`r leaf_temp_diff`}{\degreeCelsius}$ warmer on average ($n = `r n_leaf_temp`$ leaves).

### Leaf trait measurements

We selected a fully expanded, unshaded leaf at least six leaves above the cotyledons during early vegetative growth. This typically meant that plants had grown in light treatments for $\approx$ 4 weeks, ensuring they had time to sense and respond developmentally to the light intensity of the treatment rather than the seedling conditions [@schoch_dependence_1980]. Shade plants grew slower than sun plants, hence leaves at the same developmental stage were measured on chronologically older plants in the shade treatment. In some sun plants, we had to use leaves higher on the stem because short internodes made lower leaves inaccessible with the gas exchange equipment. We measured terminal leaflets in `r percent_terminal`\% of cases, but used the lateral leaflet closest to the terminal leaflet when it was damaged or difficult to clamp into the gas exchange chamber. When a leaflet was damaged during gas exchange measurements, we collected anatomical data from the nearest leaflet on the same leaf (`r percent_change_leaflet`\% of leaves).

#### Amphistomy advantage

We estimated 'amphistomy advantage' (\aax) *sensu* @parkhurst_adaptive_1978, but with modifications previously described in @triplett_amphistomy_2024. \aax{} is calculated as the log-response ratio of $A$ compared at the same total \gsw:

$$\mathrm{AA} = \mathrm{log}(A_{\mathrm{amphi}} / A_{\mathrm{hypo}})$$

We measured the photosynthetic rate of an untreated amphistomatous leaf (\Aamphi) over a range of \gsw{} values. We refer to this as an \agcurve{} curve. We compared the \agcurve{} curve of the untreated leaf to the photosynthetic rate of pseudohypostomatous leaf (\Ahypo), which is the same leaf but with gas exchange through the upper surface blocked by a neutral density plastic (propafilm).

We measured \agcurve{} curves using a portable infrared gas analyzer (LI-6800PF, LI-COR Biosciences, Lincoln, Nebraska, USA). Light-acclimated plants were placed under LEDs dimmed to match their light treatment during gas exchange measurements. We estimated the photosynthetic rate ($A$) and stomatal conductance to CO$_2$ (\gsw) at ambient CO$_2$ (\caequals{415}) and \tleafequals{25.0}. The irradiance of the light source in the pseudohypo leaf was higher because the propafilm reduces transmission. To compensate for reduced transmission, we increased incident \ppfd{} for pseudohypo leaves by a factor 1/0.91, the inverse of the measured transmissivity of the propafilm. We also set the stomatal conductance ratio, for purposes of calculating boundary layer conductance, to 0 for pseudohypo leaves following manufacturer directions.

We collected four \agcurve{} curves per leaf, an amphi (untreated) curve and a pseudohypo (treated) curve at high light-intensity (\ppfdequals{2000}; `r p_red_high`:`r p_blue_high` red:blue) and low light-intensity (\ppfdequals{150}; `r p_red_low`:`r p_blue_low` red:blue). We always measured high light-intensity curves first because photosynthetic downregulation is faster than upregulation in these species. To control for order effects, we alternated between starting with amphi or pseudohypo leaf measurements. Unlike @triplett_amphistomy_2024, preliminary experiments with *Solanum* indicated a strong order effect in that $A$ declined in the second curve. Therefore, we made measurements over two days. On the first day, we measured high and low light-intensity curves for either amphi or pseudohypo leaves; on the second day, we measured high and low light-intensity curves on the other leaf type.

In all cases, we acclimated the focal leaf to high light (\ppfdequals{2000}) and high relative humidity (\rhequals{70}) until $A$ and \gsw{} reach their maximum. After that, we decreased \rh{} to $\approx 10\%$ to induce rapid stomatal closure without biochemical downregulation. Hence, \Aamphi{} and \Ahypo{} were both measured at low chamber humidity after the leaf had acclimated to high humidity. All other environmental conditions in the leaf chamber remained the same. We logged data until \gsw{} reached its nadir. We then acclimated the leaf to low light (\ppfdequals{150}) and \rhequals{70} before inducing stomatal closure with low \rh{} and logging data as described above.

#### Stomatal anatomy

We estimated the stomatal density and size on ab- and adaxial leaf surfaces from all leaves, using guard cell length as a proxy for stomatal size since it is proportional to maximum conductance [@sack_developmental_2016]. We made surface impressions of leaf lamina from the same area used for gas exchange measurements using a-silicone impression material (Zhermack elite HD+, light body, fast set, Rovigo, Italy). We applied clear nail polish to make positive replicas of the impression. After nail polish dried, we mounted replicas on a microscope slide using transparent tape [@mott_amphistomy_1991]. We digitized a portion of each leaf surface replica using a brightfield microscope (Leica DM2000, Wetzlar, Germany). We counted and measured guard cell length on all stomata using the FIJI implementation of ImageJ2 version 2.3.0 [@schindelin_fiji_2012], then divided the count by the visible leaf area ($\qty{0.890}{\raiseto{2}\milli\meter}$) to estimate stomatal density. For each surface we calculated the anatomical maximum stomatal conductance (\gmax) at a reference leaf temperature of $\qty{25}{\degreeCelsius}$ following @sack_developmental_2016 as:

$$g_\mathrm{max} = b m D_\mathrm{S} A_\mathrm{S}.$$

The biophysical and morphological constants $b$ and $m$ are:

$$b = \frac{D_\mathrm{wv}}{v},~\text{and}$$	
$$m = \frac{\pi c ^ 2}{j^{0.5} (4 h j + \pi c)},$$

where $D_\mathrm{wv}$ is the diffusion coefficient of water vapor in air and $v$ is by the kinematic viscosity of dry air. We assumed $D_\mathrm{wv} = \qty{2.49e-5}{\meter\squared\per\second}$ and $v = \qty{2.24e-2}{\raiseto{3}\meter\per\mol}$ [@monteith_principles_2013]. For kidney-shaped guard cells like those in wild tomatoes, $c = h = j = 0.5$. The \gmaxratio{} is the \gmax{} of the adaxial surface divided by the sum total \gmax{} of both surfaces.

#### Leaf mass per area

Leaf mass per area (LMA) is the dry mass divided by the leaflet area. We scanned fresh leaflets on a flat bed scanner (Epson V600, Los Alamitos, California. USA) and measured leaflet area from digital images using the FIJI implementation of ImageJ2 version 2.3.0 [@schindelin_fiji_2012]. We dried leaves for 72 hours at $\qty{74}{\degreeCelsius}$ in a food dehydrator (Cosori CP267-FD, Vesync Co., Anaheim, California, USA) and weighed using a benchtop analytical balance (Ohaus PR64 Analytical Balance, Parsippany, New Jersey, USA). In $`r percent_lma_switch`\%$ of leaves we measured LMA on the adjacent leaflet because the focal leaflet was damaged or wilted while making surface impressions and we could not reliably estimate area. LMA data are missing from $`r percent_lma_missing`\%$ of individuals because the area or mass was not recorded at all or recorded incorrectly.

### Cleaning \agcurve{} curves

The raw data set consisted of `r prettyNum(aa_stats$n_rh_curve0, big.mark = ",")` \agcurve{} curves with an average of `r formatC(aa_stats$n_point_per_rh_curve0, digits = 3)` points per curve. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into seven sequential steps (\autoref{tbl-cleaning1}).

```{r, cleaning1, echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-cleaning1
#| tbl-cap: "Seven sequential steps for cleaning \\agcurve{} curves. The rationale and procedure for each step are described in the text. The rightmost columns summarize the number of curves and mean number of points per curve remaining after each step. For reference, there are four possible \\agcurve{} curves per replicate: all combinations of leaf type (amphi or pseudohypo) and light intensity (high or low)."

step_description_col = c(
  "1. remove unreliable and unusable data points",
  "2. remove points not at steady state",
  "3. remove hysteretic portion of \\agcurve{} curves at low \\gsw{}",
  "4. remove outliers within each \\agcurve{} curve",
  "5. remove replicates with no overlap between amphi and pseudohypo \\agcurve{} curves",
  "6. thin redundant data points within each \\agcurve{} curve",
  "7. trim extreme \\aax{} values"
)

n_curve_col = unlist(aa_stats[str_detect(names(aa_stats), "^n_rh_curve[1-9]$")]) |>
  prettyNum(big.mark = ",")

n_point_col = unlist(aa_stats[str_detect(names(aa_stats), "^n_point_per_rh_curve[1-9]$")])

tibble(
  `Step: description` = step_description_col,
  `Number of curves` = n_curve_col,
  `Number of points per curve` = n_point_col
) |>
  knitr::kable(format = "latex",
               align = c("l", "c", "c"),
               digits = c(0, 0, 1),
               booktabs = TRUE, escape = FALSE) |>
  kableExtra::kable_styling(latex_options = c("striped", "no_vertical_lines")) |>
  kableExtra::column_spec(1, width = "9cm") |>
  kableExtra::column_spec(2, width = "2.5cm") |>
  kableExtra::column_spec(3, width = "2.5cm")

```

#### Remove points not at steady state

*Rationale*: Recent methods have been developed to quantify dynamic (i.e. unsteady state) assimilation and transpiration rates during gas exchange measurements [@saathoff_gas_2021]. These methods potentially obviate the need to wait minutes to assure steady state conditions, enabling rapid estimation of dynamic gas exchange. The equations for dynamic assimilation and transpiration can also be used to assess whether steady state conditions have been reached. If dynamic and steady state estimates are similar, this indicates that the system is in steady state [@saathoff_gas_2021]. To be conservative, we did not use dynamic assimilation or transpiration rates, but instead used the comparison between dynamic and steady state rates to identify and remove points that were not at steady state.

*Procedure*: We removed all points where $\left| \log{A_\mathrm{sty} / A_\mathrm{dyn}} \right| < `r aa_args[['steadystate_cutoff']]`$ and $\left| \log{E_\mathrm{sty} / E_\mathrm{dyn}} \right| < `r aa_args[['steadystate_cutoff']]`$. The subscripts 'sty' and 'dyn' refer to steady state and dynamic estimates, respectively; the assimilation and transpiration rates are denoted by $A$ and $E$, respectively.

#### Remove unreliable and unusable data points

*Rationale*: Unreliable data points consisted of those where chamber [CO$_2$] was unstable and therefore measurements are not biologically meaningful. Unusable data points were those where $A < 0$ because the logarithm of a negative number is undefined.

*Procedure*: We retained data points where \cabetween{`r ca_min`}{`r ca_max`} and $A> 0$.

#### Remove hysteretic portion of \agcurve{} curves at low \gsw{}

*Rationale*: In most \agcurve{} curves, we observed a hysteretic response at low \gsw. After \gsw{} and $A$ declined simultaneously, $A$ increased slightly as \gsw{} continued to decline or stabilize, indicating some leaf acclimation to low \rh. We removed this portion of the curve to focus curve-fitting on the primary domain where $A$ increases monotonically with \gsw{}.

*Procedure*: For each curve, we removed data points after \gsw{} had reached its minimum unless there were fewer than `r n_min_points_per_rh_curve` data points remaining.

#### Remove outliers within each \agcurve{} curve

*Rationale*: Individual outliers within \agcurve{} curves, usually caused by transitory changes in chamber conditions, exert undue leverage on parameter estimates and cause bias and/or low precision in parameter estimates.

*Procedure*: We fit provisional quadratic regressions for each curve using ordinary least squares with the `lm()` function in *R*. We sequentially removed data points with an absolute external studentized residual $> `r point_outlier_threshold`$ until none remained.

#### Thin redundant data points within each \agcurve{} curve

*Rationale*: Data points closely spaced along the \agcurve{} curve provide redundant information and may be highly correlated (i.e. pseudoreplication). This occurred because data was logged at a constant temporal interval, but the rate at which \gsw{} declined was not constant. Thinning reduces parameter estimation bias toward densely sampled regions of the curve which may not be the most biologically informative.

*Procedure*: We retained the maxima and minima \gsw{} for each curve and thinned all but one point per thinning interval of $`r thinning_interval`~\log(\si{\mol\raiseto{-2}\meter\raiseto{-1}\second})$, retaining the point nearest the midpoint of the interval.

#### Remove replicates with no overlap between amphi and pseudohypo \agcurve{} curves

*Rationale*: We could not estimate \aax{} for replicates where amphi and pseudohypo \agcurve{} curves did not overlap.

*Procedure*: We removed replicates where the range of \gsw{} values for amphi and pseudohypo \agcurve{} curves did not overlap.

#### Trim extreme \aax{} values

*Rationale*: Extreme \aax{} values were likely due to measurement error or leaf damage. Since amphi and pseudohypo \agcurve{} curves are measured on consecutive days, a poor calibration or a damaged leaf could cause a large difference in $A$ between days, which would appear as an extreme \aax{} value.

*Procedure*: We provisionally estimated \aax{} for each replicate by integrating over the range of \gsw{} values where amphi and pseudohypo \agcurve{} curves overlap. In this procedure, curve parameters were provisionally estimated using ordinary least squares with the `lm()` function in *R*. We then used point estimates of \aax{} for each replicate as the response variable in a linear model with light treatment, light intensity, population, and all interactions as explanatory variables. This model was also fit using ordinary least squares with the `lm()` function in *R*. We classified  extreme \aax{} values as those with an absolute internal studentized residual $> `r aa_outlier_threshold`$. Because these values likely indicate significant measurement error or leaf damage, we removed \agcurve{} curves at both light intensities if either was classified as extreme.

### Bayesian data analysis with *Stan* {#sec-analysis}

Our analysis pipeline consisted of four main steps:

1. Estimate \agcurve{} curve parameters
2. Estimate \aax{} for both measurement light intensities using \agcurve{} curve parameters
3. Estimate the effects of light intensity, light treatment, and population on \aax{} (acclimation and plasticity hypotheses)
4. Estimate the effects of native light habitat on population-level \aax{} (constitutive hypothesis)

All models were fit using a Bayesian HMC sampling in the probabilistic programming language *Stan* [@stan_development_team_stan_2025] using the *R* package **brms** version `r fit_aa1$version$brms` [@burkner_brms_2017]. We used *CmdStan* version `r fit_aa1$version$cmdstan` and **cmdstanr** version `r fit_aa1$version$cmdstanr` [@gabry_cmdstanr_2025] to interface with *R* version `r getRversion()` [@r_core_team_r:_2025]. We sampled the posterior distribution from a single chain with a minimum of `r aa_args$n_iter_init / 2` iterations after `r aa_args$n_iter_init / 2` warmup iterations. If necessary, we refit models with more iterations to decrease the Gelman-Rubin convergence statistic ($\hat{R}$) [@gelman_inference_1992] to less than `r aa_args[["max_rhat"]]` and the bulk effective sample size (ESS) to greater than `r aa_args[["min_ess"]]` for all model parameters, and there were fewer than `r aa_args[["max_divergent"]]` divergent transitions.

#### \agcurve{} curve parameters {#sec-agcurve-pars}

We modeled \logA{} as a quadratic function of \loggsw{} for each leaf, measured at low and high light intensity, in both amphistomatous (untreated) and pseudohypostomatous (propafilm treatment) states.

#### \aax{} for each light intensity within leaf using \agcurve{} curve parameters {#sec-aa-est}

We estimated \aax{} for each light intensity by integrating the difference in \logA{} between the amphi and pseudohypo \agcurve{} curves over the range of \gsw{} values where the curves overlap (from $\min(\log(g_\text{sw}))$ to $\max(\log(g_\text{sw}))$).

$$\widehat{\mathrm{AA}} = \int_{\text{min(log(}g_\text{sw}))}^{\text{max(log(}g_\text{sw}))} \text{log}\bigg(\frac{\hat{A}_\text{amphi}(x; \theta_{\text{amphi}})}{\hat{A}_\text{hypo}(x; \theta_{\text{hypo}})}\bigg) dx$$
where $\theta_\text{amphi}$ and $\theta_\text{hypo}$ are the quadratic parameters of the amphi and pseudohypo \agcurve{} curves, respectively (@fig-example-aa). We calculated $\widehat{\mathrm{AA}}$ from each draw of the posterior for the paired amphi and pseudohypo \agcurve{} curves to obtain the posterior distribution of $\widehat{\mathrm{AA}}$ for each leaf at each light intensity. We then summarized the posterior distribution of $\widehat{\mathrm{AA}}$ for each leaf at each light intensity by the posterior median and standard deviation, which became our point estimate of, and uncertainty in, $\widehat{\mathrm{AA}}$ for the phylogenetic mixed effects model described below.

```{r, example-aa, echo=FALSE,message=FALSE,warning=FALSE}
#| label: fig-example-aa
#| fig-align: center
#| fig-width: 6.5
#| fig-height: 3.5
#| fig-cap: "Example amphistomy advantage (\\aax) calculation from \\agcurve{} curves in an amphistomatous leaf with gas exchange through both surfaces (blue) and the same leaf with gas exchange blocked on the upper surface only (orange). We integrate over the gray region between the curves to estimate \\aax{}. The \\aax{} is low (left facet) when the photosynthetic rate of an amphistomatous leaf is similar to a pseudohypostomatous leaf at the same total stomatal conductance ($x$-axes); large \\aax{} (right facet) is indicated when an amphistomatous leaf has a higher photosynthetic rate than a pseudohypostomatous leaf."
read_rds("../figures/example-aa.rds")
```

#### Phylogenetic mixed effects models {#sec-phylo}

We fit Bayesian mixed effects models with phylogenetically structured random effects to test predictions of competing hypotheses about why amphistomy advantage (\aax) might be greater for leaves in sunny, open habitats. We fit all models combinations of measurement light intensity, growth light intensity, and their interaction as both fixed effects and random effects among populations. In other words, we test whether there were main effects of light intensity treatments and/or whether populations varied in their response to light intensity treatments. In all models, we included population as a phylogenetically structured random effect. We accounted for uncertainty in $\widehat{\mathrm{AA}}$ using the standard deviation of the posterior distribution of each $\widehat{\mathrm{AA}}$ estimate (see above). We crossed fixed and random effect structures with fixed effects of measurement light intensity and/or growth light intensity on the distributional parameter, $\sigma$, the residual variance in \aax{}. We used a robust regression approach by assuming $t$-distributed residuals, which makes estimates less susceptible to extreme values. In total, we fit `r n_model1` models, each with a different combination of fixed, random, and distributional effects. We used the RAxML whole-transcriptome concatenated phylogeny based on 2,745 100-kb genomic windows from @pease_phylogenomics_2016. Two of our populations were not in this tree. We used accession LA1044 in place of LA3909, two populations of *S. galapagense*; LA0750 was added as sister to LA0716, two closely related populations of *S. pennellii*. The node separating LA0750 from LA0716 was placed half-way between the next deepest node. We used the leave-one-out cross-validation information criterion (LOOIC) to compare the fit of models (\autoref{tbl-aa_loo1}) using the *R* package **loo** version `r packageVersion("loo")` [@vehtari_practical_2017] to calculate LOOIC values. We selected the model with the lowest LOOIC value (@tbl-aa_loo1) to generate posterior predictions for hypothesis testing. Parameter estimates from the selected model are shown in @tbl-fit_aa1.

To test whether LMA mediated the relationship between growth light intensity and \aax{}, we repeated a similar model comparison procedure but included $\log \mathrm{LMA}$ as an additional fixed effect in the model, as well its interaction with measurement and growth light intensity. The model included a direct effect of growth light intensity on $\log \mathrm{LMA}$ and random phylogeneticlly structured effects of population on $\log \mathrm{LMA}$ and effects of sun treatment on $\log \mathrm{LMA}$. This enabled us to disentangle whether $\log \mathrm{LMA}$ effects \aax{} directly or indirectly because both traits respond to growth light intensity. We integrated over missing LMA data using the `mi()` function in **brms**. Because the evidence from the first set of models strongly supported including effects of growth and measurement light intensity on the residual variance in $\mathrm{AA}$, we included these in terms in all models. In total, we fit `r n_model2` models, each with a different combination of fixed and random effect structures. We selected the model with the lowest LOOIC value (@tbl-aa_loo2) to generate posterior predictions for hypothesis testing. Note that LOOIC was calculated only from pointwise likelihood values of the submodel estimating effects of light intensity, light treatments, and population on \aax{}. 
Parameter estimates from the selected model are shown in @tbl-fit_aa2.

#### Posterior predictions {#sec-predictions}

The acclimation, plasticity, and constitutive hypotheses make different predictions about the relationship between \aax{} and light intensity, light treatment, and native $\mathrm{PAI}$ among populations. Since these hypotheses are not mutually exclusive, we describe how we assessed support for one, two, or all three hypotheses simultaneously in \autoref{tbl-predictions}. In general, the acclimation hypothesis was supported if \aax{} was greater at high light intensity than low light intensity. The plastic hypothesis was supported if \aax{} was greater in sun leaves than shade leaves. The constitutive hypothesis was supported if population-level \aax{} increased with native \ppfd. In interactive models, we only consider positively reinforcing interactions between high light intensity, sun leaves, and native \ppfd{} because these are the only interactions which could explain why amphistomatous leaves are advantageous in high light habitats.

From the model selected using LOOIC, we predicted the posterior distribution of the expected \aax{} for each population at each measurement light intensity and growth light intensity. If the 95% confidence intervals for \aax{} did not overlap zero, we considered \aax{} to be significantly different from zero. We also calculated posterior distribution for the average \aax{} among populations in each condition. We tested a linear effect of native light habitat on population-level \aax{} in each treatment condition, integrating over the entire posterior distribution to calculate a posterior distribution of the slope. If the 95\% confidence interval for the slope did not overlap zero, we considered the slope to be significant.

```{r, predictions, eval=TRUE, echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-predictions
#| tbl-cap: "Predictions of competing hypotheses about the relationship between \\aax{} and light intensity, light treatment, and native plant area index (PAI) among populations. The middle column lists specific directional predictions about parameter values. The rightmost column describes the predictions in words and explains how population-level \\aax{} is calculated in the relevant model."

tribble(
  ~"\\textbf{Hypothesis}", ~"\\textbf{Prediction(s)}", ~"\\textbf{Description}",
  "Acclimation", "$\\mathrm{AA}_{2000} > 0$", "\\aax{} at high light intensity is greater than that at low light intensity",
  "Plasticity", "$\\mathrm{AA}_{\\text{sun}} > 0$", "\\aax{} in sun leaves is greater than that in shade leaves",
  "Constitutive", "$\\beta_{\\mathrm{PAI,AA}} < 0$", "Population-level \\aax{} ($\\mathrm{AA}_\\text{pop}$) decreases with native PAI",
  "Constitutive", "", "$\\mathrm{AA}_\\text{pop} = \\beta_{\\mathrm{AA}, 0} + \\vec{\\beta}_{\\mathrm{AA}, \\text{pop}}$",
  "Acclimation $\\times$ Plasticity", "$\\mathrm{AA}_{2000,\\text{sun}} > 0$", "\\aax{} is highest at high light intensity in sun leaves",
  "Acclimation $\\times$ Plasticity", "", "",
  "Acclimation $\\times$ Constitutive", "$\\mathrm{AA}_{2000} > 0$", "\\aax{} at high light intensity is greater than that at low light intensity",
  "Acclimation $\\times$ Constitutive", "$\\beta_{\\mathrm{PAI,AA}} < 0$", "Population-level \\aax{} at high light intensity ($\\mathrm{AA}_{\\text{pop},2000}$) decreases with native PAI",
  "Acclimation $\\times$ Constitutive", "", "$\\mathrm{AA}_{\\text{pop},2000} = \\beta_{\\mathrm{AA}, 0} + \\vec{\\beta}_{\\mathrm{AA}, \\text{pop}} + \\vec{\\beta}_{\\mathrm{AA}, 2000, \\text{pop}}$",
  "Plasticity $\\times$ Constitutive", "$\\mathrm{AA}_{\\text{sun}} > 0$", "\\aax{} in sun leaves is greater than that in shade leaves",
  "Plasticity $\\times$ Constitutive", "$\\beta_{\\mathrm{PAI,AA}} < 0$", "Population-level \\aax{} in sun leaves ($\\mathrm{AA}_{\\text{pop},\\text{sun}}$) decreases with native PAI",
  "Plasticity $\\times$ Constitutive", "", "$\\mathrm{AA}_{\\text{pop},\\text{sun}} = \\beta_{\\mathrm{AA}, 0} + \\vec{\\beta}_{\\mathrm{AA}, \\text{pop}} + \\vec{\\beta}_{\\mathrm{AA}, \\text{sun}, \\text{pop}}$",
  "Acclimation $\\times$ Plasticity $\\times$ Constitutive", "$\\mathrm{AA}_{2000,\\text{sun}} > 0$", "\\aax{} is highest at high light intensity in sun leaves",
  "Acclimation $\\times$ Plasticity $\\times$ Constitutive", "$\\beta_{\\mathrm{PAI,AA}} > 0$", "Population-level \\aax{} at high light intensity in sun leaves ($\\mathrm{AA}_{\\text{pop},2000,\\text{sun}}$) decreases with native PAI",
  "Acclimation $\\times$ Plasticity $\\times$ Constitutive", "", "$\\mathrm{AA}_{\\text{pop},2000,\\text{sun}} = \\beta_{\\mathrm{AA}, 0} + \\vec{\\beta}_{\\mathrm{AA}, \\text{pop}} + \\vec{\\beta}_{\\mathrm{AA}, 2000, \\text{pop}} + \\vec{\\beta}_{\\mathrm{AA}, \\text{sun}, \\text{pop}}$"

) |>
  knitr::kable(
    format = "latex",
    align = c("l", "l", "l"),
    booktabs = TRUE,
    escape = FALSE,
    longtable = TRUE
  ) |>
  kableExtra::add_indent(positions = seq_len(15), level_of_indent = -1, target_cols = 3) |>
  kableExtra::collapse_rows(1, latex_hline = "major", valign = "top",
                            row_group_label_position = "first") |>
  kableExtra::column_spec(1, width = "1in") |>
  kableExtra::column_spec(2, width = "1.5in") |>
  kableExtra::column_spec(3, width = "3in")

```

### The water cost of hypostomy

Amphistomy advantage implies that hypostomatous leaves would need to lose more water by having higher \gsw{} to achieve the same photosynthetic rate as an otherwise identical amphistomatous leaf. We estimated this cost of hypostomy approximately. For simplicity, we focus on a sufficiently small region of the \agcurve{} curve such that we can approximate the relationship between \gsw{} and $A$ as log-log linear. We also assume that the slope is the same in both amphistomatous and hypostomatous leaves, which was often approximately the case in our experiment (@fig-curves). With these assumptions, the approximate water cost of hypostomy can be derived from the equations:

$$\log(A_{\mathrm{amphi}}) = \beta_\text{amphi} + \varepsilon_g \log(g_\text{sw}),~\text{and}$$
$$\log(A_{\mathrm{hypo}}) = \beta_\text{hypo} + \varepsilon_g \log(g_\text{sw}).$$
Solving for the difference in $\log{g_\text{sw}}$ at a given $\log{A_{\mathrm{hypo}}}$ is:

$$\log(g_\text{sw,amphi}) - \log(g_\text{sw,hypo}) = \frac{\beta_\text{amphi} - \beta_\text{hypo}}{\varepsilon_g}.$$
Noting that because the curves are parallel, $\mathrm{AA} = \log(A_\text{amphi} /A_\text{hypo}) = \beta_\text{amphi} - \beta_\text{hypo}$, this reveals that the negative log-ratio of the \gsw{} is related to \aax{} divided by the slope:

$$-\log\left(\frac{g_\text{sw,amphi}}{g_\text{sw,hypo}}\right) = \frac{\mathrm{AA}}{\varepsilon_g}.$$
We note that the slope in this case is the elasticity of $A$ to \gsw{} because both are log-transformed. This quantity will be proportional the water cost of hypostomy assuming that VPD and other environmental variables are held constant and the boundary layer resistance is negligible.

<!-- estimated the average value of this quantity for each population under each measurement light treatment by regressing \gsw{} against $A$ on a log-log scale, with curve type (amphistomy versus pseudohypostomy) and replicate within population as a random effect using **brms**. We followed the fitting procedures described above for other models. Because of the different assumptions, the estimates of \aax{} in this model will not be identical to that described above, but this approximate method enables a simpler, back-of-the-envelope calculation of the water cost of hypostomy. -->

```{r fit_aa1, results = 'asis', echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
#| label: tbl-fit_aa1
#| tbl-cap: "Parameter estimates and 95% confidence intervals (CIs) from the posterior distribution of the selected model. Models potentially include fixed, random, and distributional effects of growth light intensity (shade and sun), measurement light intensity (low and high), and their interaction on amphistomy advantage ($\\mathrm{AA}$). The distributional effects refer to the effect of factors on residual variation in $\\mathrm{AA}$, denoted $\\sigma$, on a log-link scale. For phylogenetically structured random effects, we report the estimated standard deviation (SD)."

make_brms_summary_table = function(fit,
                                   ci_level = 0.95,
                                   digits = 2) {
  # Get posterior summary
  summ = posterior_summary(fit, prob = c((1 - ci_level) / 2, 1 - (1 - ci_level) / 2))
  summ_df = as.data.frame(summ)
  summ_df$param = rownames(summ_df)
  colnames(summ_df) = c("Estimate", "Est.Error", "CI_low", "CI_high", "Parameter")

  # Classify parameters
  summ_df = summ_df |>
    mutate(
      Type = case_when(
        grepl("^b_", Parameter) &
          !grepl("sigma|Intercept", Parameter) ~ "fixed effects",
        grepl("^b_Intercept$", Parameter) ~ "fixed effects",
        grepl("^b_", Parameter) &
          grepl("sigma", Parameter) ~ "distributional parameters",
        grepl("^sd_", Parameter) ~ "random effect SDs",
        TRUE ~ "Other"
      )
    )

  # Clean parameter names
  summ_df = summ_df |>
    mutate(
      Parameter = case_when(
        Parameter == "b_Intercept" ~ "$\\mathrm{AA}$ intercept (shade, low light)",
        Parameter == "b_light_treatmenthigh" ~ "effect of sun on $\\mathrm{AA}$",
        Parameter == "b_light_intensity2000" ~ "effect of high light on $\\mathrm{AA}$",
        Parameter == "b_light_treatmenthigh:light_intensity2000" ~ "effect of sun $\\times$ high light interaction on $\\mathrm{AA}$",
        Parameter == "b_sigma_Intercept" ~ "$\\log \\sigma$ intercept (shade, low light)",
        Parameter == "b_sigma_light_treatmenthigh" ~ "effect of sun on $\\log \\sigma$",
        Parameter == "b_sigma_light_intensity2000" ~ "effect of high light on $\\log \\sigma$",
        Parameter == "sd_acc__Intercept" ~ "$\\mathrm{AA}$ among populations",
        Parameter == "sd_acc__light_treatmenthigh" ~ "effect of sun on $\\mathrm{AA}$ among populations",
        Parameter == "sd_acc__light_intensity2000" ~ "effect of high light on $\\mathrm{AA}$ among populations",
        Parameter == "sd_acc__light_treatmenthigh:light_intensity2000" ~ "effect of sun $\\times$ high light interaction on $\\mathrm{AA}$ among populations"
      )
    )

  # Select and format
  summ_df |>
    filter(Type != "Other") |>
    mutate(across(c(Estimate, CI_low, CI_high), ~ round(., digits))) |>
    mutate(`Estimate [95% CI]` = paste0(Estimate, " [", CI_low, ", ", CI_high, "]")) |>
    select(Type, Parameter, `Estimate [95% CI]`) |>
    arrange(match(
      Type,
      c(
        "fixed effects",
        "distributional parameters",
        "random effect SDs"
      )
    )) |>
    as_tibble()
}

summary_tbl = make_brms_summary_table(fit_aa1)

tbl_fixed = summary_tbl |> filter(Type == "fixed effects") |> select(-Type)
tbl_dist  = summary_tbl |> filter(Type == "distributional parameters") |> select(-Type)
tbl_rand  = summary_tbl |> filter(Type == "random effect SDs") |> select(-Type)

bind_rows(tbl_fixed, tbl_dist, tbl_rand) |>
  kable(
    col.names = c("Parameter", "Estimate [95\\% CI]"),
    booktabs = TRUE,
    format = "latex",
    escape = FALSE
  ) |>
  kable_styling(latex_options = c("striped")) |>
  pack_rows("Fixed effects",
            1,
            nrow(tbl_fixed),
            bold = FALSE,
            italic = TRUE) |>
  pack_rows(
    "Distributional parameters",
    nrow(tbl_fixed) + 1,
    nrow(tbl_fixed) + nrow(tbl_dist),
    bold = FALSE,
    italic = TRUE
  ) |>
  pack_rows(
    "Random effect SDs",
    nrow(tbl_fixed) + nrow(tbl_dist) + 1,
    nrow(tbl_fixed) + nrow(tbl_dist) + nrow(tbl_rand),
    bold = FALSE,
    italic = TRUE
  ) |>
  column_spec(1, width = "4in") |>
  column_spec(2, width = "2in")

```

```{r fit_aa2, results = 'asis', echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
#| label: tbl-fit_aa2
#| tbl-cap: "Parameter estimates and 95% confidence intervals (CIs) from the posterior distribution of the selected model with leaf mass per area (LMA) as potential mediator variable. Models potentially include fixed and random effects of growth light intensity (shade and sun), measurement light intensity (low and high), $\\log \\mathrm{LMA}$, and their interactions on amphistomy advantage ($\\mathrm{AA}$). LMA can also be affected by growth light intensity and population. For phylogenetically structured random effects, we report the estimated standard deviation (SD)."

make_brms_summary_table = function(fit,
                                   ci_level = 0.95,
                                   digits = 2) {
  # Get posterior summary
  summ = posterior_summary(fit, prob = c((1 - ci_level) / 2, 1 - (1 - ci_level) / 2))
  summ_df = as.data.frame(summ)
  summ_df$param = rownames(summ_df)
  colnames(summ_df) = c("Estimate", "Est.Error", "CI_low", "CI_high", "Parameter")

  # Classify parameters
  summ_df = summ_df |>
    mutate(
      Type = case_when(
        grepl("^b_", Parameter) &
          !grepl("sigma|Intercept", Parameter) ~ "fixed effects",
        grepl("^b_(aa|llma)_Intercept$", Parameter) ~ "fixed effects",
        grepl("^bsp_", Parameter) ~ "fixed effects",
        grepl("^b_", Parameter) &
          grepl("sigma", Parameter) ~ "distributional parameters",
        grepl("^sd_", Parameter) ~ "random effect SDs",
        TRUE ~ "Other",
      ),
      resp = case_when(
        str_detect(Parameter, "_aa_") ~ "aa",
        str_detect(Parameter, "_llma_") ~ "lma",
        TRUE ~ NA_character_
      )
    )

  # Clean parameter names
  summ_df <- summ_df %>%
    mutate(
      Parameter = case_when(
        # AA
        Parameter == "b_aa_Intercept" ~ "$\\mathrm{AA}$ intercept (shade, low light)",
        Parameter == "b_aa_light_treatmenthigh" ~ "effect of sun on $\\mathrm{AA}$",
        Parameter == "b_aa_light_intensity2000" ~ "effect of high light on $\\mathrm{AA}$",
        Parameter == "b_aa_light_treatmenthigh:light_intensity2000" ~ "effect of sun $\\times$ high light interaction on $\\mathrm{AA}$",
        Parameter == "bsp_aa_millma" ~ "effect of $\\log \\mathrm{LMA}$ on $\\mathrm{AA}$",
        Parameter == "b_sigma_aa_Intercept" ~ "$\\log \\sigma$ intercept (shade, low light)",
        Parameter == "b_sigma_aa_light_treatmenthigh" ~ "effect of sun on $\\log \\sigma$",
        Parameter == "b_sigma_aa_light_intensity2000" ~ "effect of high light on $\\log \\sigma$",
        Parameter == "sd_acc__aa_Intercept" ~ "$\\mathrm{AA}$ among populations",
        Parameter == "sd_acc__aa_light_treatmenthigh" ~ "effect of sun on $\\mathrm{AA}$ among populations",
        Parameter == "sd_acc__aa_light_intensity2000" ~ "effect of high light on $\\mathrm{AA}$ among populations",
        Parameter == "sd_acc__aa_light_treatmenthigh:light_intensity2000" ~ "effect of sun $\\times$ high light interaction on $\\mathrm{AA}$ among populations",

        # LMA
        Parameter == "b_llma_Intercept" ~ "$\\log \\mathrm{LMA}$ intercept (shade)",
        Parameter == "b_llma_light_treatmenthigh" ~ "effect of sun on $\\log \\mathrm{LMA}$",
        Parameter == "b_sigma_llma_Intercept" ~ "$\\log \\sigma$ intercept (shade)",
        Parameter == "b_sigma_llma_light_treatmenthigh" ~ "effect of sun on $\\log \\sigma$",
        Parameter == "sd_acc__llma_Intercept" ~ "$\\log \\mathrm{LMA}$ among populations",
        Parameter == "sd_acc__llma_light_treatmenthigh" ~ "effect of sun on $\\log \\mathrm{LMA}$ among populations"
      )
    )

  # Select and format
  summ_df |>
    filter(Type != "Other") |>
    mutate(across(c(Estimate, CI_low, CI_high), ~ round(., digits))) |>
    mutate(`Estimate [95% CI]` = paste0(Estimate, " [", CI_low, ", ", CI_high, "]")) |>
    select(resp, Type, Parameter, `Estimate [95% CI]`) |>
    arrange(resp, match(
      Type,
      c(
        "fixed effects",
        "distributional parameters",
        "random effect SDs"
      )
    )) |>
    as_tibble()
}

summary_tbl = make_brms_summary_table(fit_aa2)

tbl_aa_fixed = summary_tbl |> filter(Type == "fixed effects", resp == "aa") |> select(-Type)
tbl_aa_dist  = summary_tbl |> filter(Type == "distributional parameters", resp == "aa") |> select(-Type)
tbl_aa_rand  = summary_tbl |> filter(Type == "random effect SDs", resp == "aa") |> select(-Type)
tbl_lma_fixed = summary_tbl |> filter(Type == "fixed effects", resp == "lma") |> select(-Type)
tbl_lma_rand  = summary_tbl |> filter(Type == "random effect SDs", resp == "lma") |> select(-Type)

bind_rows(tbl_aa_fixed,
          tbl_aa_dist,
          tbl_aa_rand,
          tbl_lma_fixed,
          tbl_lma_rand) |>
  select(-resp) |>
  kable(
    col.names = c("Parameter", "Estimate [95\\% CI]"),
    booktabs = TRUE,
    format = "latex",
    escape = FALSE
  ) |>
  kable_styling(latex_options = c("striped")) |>
  pack_rows(
    "Response: $\\mathrm{AA}$",
    1,
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + nrow(tbl_aa_rand),
    escape = FALSE
  ) |>
  pack_rows(
    "~~Fixed effects",
    1,
    nrow(tbl_aa_fixed),
    escape = FALSE,
    bold = FALSE,
    italic = TRUE
  ) |>
  pack_rows(
    "~~Distributional parameters",
    nrow(tbl_aa_fixed) + 1,
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist),
    escape = FALSE,
    bold = FALSE,
    italic = TRUE
  ) |>
  pack_rows(
    "~~Random effect SDs",
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + 1,
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + nrow(tbl_aa_rand),
    escape = FALSE,
    bold = FALSE,
    italic = TRUE
  ) |>
  pack_rows(
    "Response: $\\log~\\\\mathrm{LMA}$",
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + nrow(tbl_aa_rand) + 1,
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + nrow(tbl_aa_rand) + nrow(tbl_lma_fixed) + nrow(tbl_lma_rand),
    escape = FALSE
  ) |>
  pack_rows(
    "~~Fixed effects",
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + nrow(tbl_aa_rand) + 1,
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + nrow(tbl_aa_rand) + nrow(tbl_lma_fixed),
    escape = FALSE,
    bold = FALSE,
    italic = TRUE
  ) |>
  pack_rows(
    "~~Random effect SDs",
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + nrow(tbl_aa_rand) + nrow(tbl_lma_fixed) + 1,
    nrow(tbl_aa_fixed) + nrow(tbl_aa_dist) + nrow(tbl_aa_rand) + nrow(tbl_lma_fixed) + nrow(tbl_lma_rand),
    escape = FALSE,
    bold = FALSE,
    italic = TRUE
  ) |>
  column_spec(1, width = "4.5in") |>
  column_spec(2, width = "2in")

```

```{r, sr-rxnorm-cap, echo=FALSE,message=FALSE,warning=FALSE}

sr_rxnorm_cap = glue::glue("Developmental plasticity in anatomical maximum stomatal conductance ratio (\\gmaxratio; $y$-axis) to growth light intensity ($x$-axis) among {n_acc} wild tomato populations. All populations are amphistomatous, but vary in the stomatal density and size on each surface, which determines \\gmaxratio. Each point is the average per population in that treatment; black lines connect the same population across treatments.")

```

```{r, sr-rxnorm, echo=FALSE,message=FALSE,warning=FALSE}
#| label: fig-sr-rxnorm
#| fig-align: center
#| fig-width: 4
#| fig-height: 4
#| fig-cap: !expr sr_rxnorm_cap
read_rds("../figures/sr-rxnorm.rds")
```

```{r, aa_loo1, echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-aa_loo1
#| tbl-cap: 'Leave-one-out cross-validation information criterion (LOOIC) for models of \\aax{} as a function of light intensity, light treatment, and population. Only the top models ($\\Delta \\text{LOOIC} < 2$) are shown. The model with the lowest LOOIC was selected for generating posterior predictions to test hypotheses. The LOOIC column lists the LOOIC value for each model; the $\\Delta$LOOIC column lists the difference between the LOOIC of the corresponding model and the best model (lowest LOOIC). Checkmarks in each row indicate whether the corresponding model included affects of measurement light intensity ("Meas.") or growth light intensity ("Growth") as fixed effects, random effects, or distributional effect on the residual variance. We also test for fixed and random interaction effects ("Inter.") between measurement and growth light intensity. All models included phylogenetically structured random effects of population.'

df1 = aa_loo_table1 |>
   filter(delta_looic < 2) |>
   mutate(
    fixed_meas = str_detect(fixed, "light_intensity"),
    fixed_growth = str_detect(fixed, "light_treatment"),
    fixed_intrxn = str_detect(fixed, "\\*"),
    # random_acc = TRUE,
    random_meas = str_detect(random, "light_intensity"),
    random_growth = str_detect(random, "light_treatment"),
    random_intrxn = str_detect(random, "\\*"),
    sigma_meas = str_detect(sigma, "light_intensity"),
    sigma_growth = str_detect(sigma, "light_treatment"),
  ) |>
  dplyr::select(
    LOOIC = looic,
    `$\\Delta$LOOIC` = delta_looic,
    starts_with("fixed_"),
    starts_with("random_"),
    starts_with("sigma_")
  ) |>
  mutate(across(where(is.logical), ~ ifelse(.x, "$\\checkmark$", "")))

c1 = colnames(df1)
groupings = list(
  " " = sum(str_detect(colnames(df1), "LOOIC")),
  "Fixed" = sum(str_detect(colnames(df1), "fixed_")),
  "Random" = sum(str_detect(colnames(df1), "random_")),
  "Distributional" = sum(str_detect(colnames(df1), "sigma_"))
)

c2 = case_when(
  str_detect(c1, "_meas") ~ "Meas.",
  str_detect(c1, "_growth") ~ "Growth",
  str_detect(c1, "_intrxn") ~ "Inter.",
  str_detect(c1, "_acc") ~ "Population",
  .default = c1
)

kbl(
  df1,
  col.names = c2,
  booktabs = TRUE,
  escape = FALSE,
  align = "c",
  longtable = TRUE,
  digits = 2
) |>
  kable_styling(latex_options = c("striped", "repeat_header")) |>
  add_header_above(groupings) |>
  column_spec(1, width = "0.7in") |>
  column_spec(2, width = "0.5in") |>
  column_spec(3:10, width = "0.4in") |>
  column_spec(c(5, 8), border_right = TRUE)

```

```{r, aa_loo2, results = "asis", eval=TRUE,echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-aa_loo2
#| tbl-cap: 'Leave-one-out cross-validation information criterion (LOOIC) for models of \\aax{} as a function of light intensity, light treatment, leaf mass per area (LMA), and population. Only the top models ($\\Delta~\\text{LOOIC} < 2$) are shown. The model with the lowest LOOIC was selected for generating posterior predictions to test hypotheses. The LOOIC column lists the LOOIC value for each model; the $\\Delta$LOOIC column lists the difference between the LOOIC of the corresponding model and the best model (lowest LOOIC). Checkmarks in each row indicate whether the corresponding model included affects of measurement light intensity ("Meas."), growth light intensity ("Growth"), or LMA as fixed or random effects. We test for fixed interaction effects between measurement and growth light intensity and LMA. We also test for random interaction effects between measurement and growth light intensity. All models included phylogenetically structured random effects of population.'

df1 = aa_loo_table2 |>
  filter(delta_looic < 2) |>
  mutate(
    fixed_meas = str_detect(fixed, "light_intensity"),
    fixed_growth = str_detect(fixed, "light_treatment"),
    fixed_lma = str_detect(fixed, "mi\\(llma\\)"),
    fixed_intrxn123 = str_detect(fixed, "light_treatment \\* light_intensity \\* mi\\(llma\\)"),
    fixed_intrxn12 = str_detect(fixed, "light_treatment \\* light_intensity") | fixed_intrxn123,
    fixed_intrxn13 = str_detect(fixed, "light_treatment \\* mi\\(llma\\)") | fixed_intrxn123,
    fixed_intrxn23 = str_detect(fixed, "light_intensity \\* mi\\(llma\\)") | fixed_intrxn123,
    # random_acc = TRUE,
    random_meas = str_detect(random, "light_intensity"),
    random_growth = str_detect(random, "light_treatment"),
    random_intrxn = str_detect(random, "\\*"),
    # sigma_meas = str_detect(sigma, "light_intensity"),
    # sigma_growth = str_detect(sigma, "light_treatment"),
  ) |>
  dplyr::select(
    LOOIC = looic,
    `$\\Delta$LOOIC` = delta_looic,
    starts_with("fixed_"),
    starts_with("random_"),
    # starts_with("sigma_")
  ) |>
  mutate(across(where(is.logical), ~ ifelse(.x, "$\\checkmark$", "")))

c1 = colnames(df1)
groupings = list(
  " " = sum(str_detect(colnames(df1), "LOOIC")),
  "Fixed" = sum(str_detect(colnames(df1), "fixed_")),
  "Random" = sum(str_detect(colnames(df1), "random_"))
  # "Distributional" = sum(str_detect(colnames(df1), "sigma_"))
)

c2 = case_when(
  str_detect(c1, "_meas") ~ "Meas.",
  str_detect(c1, "_growth") ~ "Growth",
  str_detect(c1, "_lma") ~ "LMA",
  str_detect(c1, "_intrxn12") ~ "Meas. $\\times$ Growth",
  str_detect(c1, "_intrxn13") ~ "Meas. $\\times$ LMA",
  str_detect(c1, "_intrxn23") ~ "Growth $\\times$ LMA",
  str_detect(c1, "_intrxn123") ~ "Meas. $\\times$ Growth $\\times$ LMA",
  str_detect(c1, "_intrxn$") ~ "Meas. $\\times$ Growth",
  str_detect(c1, "_acc") ~ "Population",
  .default = c1
)

cat("\\begin{landscape}")

kbl(
  df1,
  col.names = c2,
  booktabs = TRUE,
  escape = FALSE,
  align = "c",
  longtable = TRUE,
  digits = 2
) |>
  kable_styling(latex_options = c("striped", "repeat_header")) |>
  add_header_above(groupings) |>
  column_spec(1, width = "0.7in") |>
  column_spec(2, width = "0.5in") |>
  column_spec(3:12, width = "0.4in")
cat("\\end{landscape}")
```

```{r, curves, echo=FALSE,message=FALSE,warning=FALSE}
#| label: fig-curves
#| fig-cap: "(Figures in separate file) The \\agcurve{} curves, fitted lines, and \\aax{} estimates for every individual plant. The title provides the Tomato Genetics Resource Center accession number, replicate letter, and species names. The subtitle indicates the growth light intensity (sun or shade). Points are raw data, lines are fitted curves, and shaded regions are 95% confidence ribbons for the fitted curves. The Bayesian correlation coefficient, $r^2$, is shown to the right of each curve, along with the estimate of \\aax{} $\\pm$ two standard errors for each pair of amphi (orange) and pseudohypo (blue) curves."
ggplot()
```